---
title: "RStudio, mlflow, and databricks"
output: html_notebook
---

```{r setup}
# Packages ----
library(sparklyr)
library(mlflow)
library(carrier)
library(tidyverse)
library(parsnip)
library(yardstick)
library(reticulate)

# Python setup ----
use_condaenv("r-mlflow")
```

```{r spark-connection}
sc <- spark_connect(method = "databricks")
```

Here are the necessary steps:
1. Register to the Databricks mlflow tracking server
2. Build models
  a. Native R model
  b. Spark model
3. Log each model as an artifact to mlflow
4. Register each model to the model registry
5. Deploy the spark model into databricks from the model registry
   - This isn't currently supported
6. Get the R model from the registry via mlflow::load_model
7. Build Plumber API around the R model
8. Build shiny app that uses both Plumber model hosted on RSC and Spark model hosted on databricks

## Setup mlflow
First, the `mlflow` R client needs to be installed and `mlflow` itself needs to be installed. Followed the instructions [here](https://github.com/mlflow/mlflow/tree/master/mlflow/R/mlflow) to get the dev version setup.

Once `mlflow` is available, followed the steps outlined here: https://docs.databricks.com/applications/mlflow/access-hosted-tracking-server.html

Created the necessary environment variables in `.Renviron`. Ensure there is no
trailing `/` in `DATABRICKS_HOST`

```
MLFLOW_TRACKING_URI=databricks
DATABRICKS_HOST=...
DATABRICKS_TOKEN=...
```

Once those values are set, the `mlflow` R package automatically connects to the
tracking server hosted by databricks.

### Specify experiment
An experiment can be connected to or created from R using
`mlflow::mlflow_set_experiment()`, which connects to an existing experiment or
creates a new one.

```{r}
(experiments <- mlflow_list_experiments())
```

```{r}
db_connect_exp_id <- experiments %>% 
  filter(str_detect(name, "db-connect")) %>% 
  pull(experiment_id)

mlflow_set_experiment(experiment_id = db_connect_exp_id)

mlflow_get_experiment()
```

## Build models
### R model
#### Data
Using `iris` here for simplicity
```{r}
iris
```

#### Define model
```{r}
model_train <- function(x, y, mtry = NULL, trees = NULL, min_n = NULL) {
 with(mlflow_start_run(), {
   # Log params to mlflow
   mlflow_log_param("mtry", mtry)
   mlflow_log_param("trees", trees)
   mlflow_log_param("min_n", min_n)
   
   # Setup model
   model <- rand_forest(mtry = mtry, trees = trees, min_n = min_n) %>% 
     set_engine("ranger") %>% 
     set_mode("classification")
   
   # Train model
   model_fit <- model %>% 
     fit_xy(x = x, y = y)
   
   # Evaluate model
   predictions <- predict(model_fit, iris) %>% 
     bind_cols(iris %>% select(Species))
   
   pred_acc <- accuracy(predictions, truth = "Species", estimate = ".pred_class")[[".estimate"]]
   
   # Log performance
   mlflow_log_metric("accuracy", pred_acc)
   
   # Create portable model with crate
   mlflow_predictor <- carrier::crate(~ranger:::predict.ranger(model$fit, .x)$predictions,
                                      model = model_fit)
   
   # Log model to mlflow
   mlflow_log_model(mlflow_predictor, "model")
 })
}
```

```{r}
labels <- iris %>% 
  select(Species)

features <- iris %>% 
  select(-Species)
```

```{r}
model_train(features, labels, mtry = 3, trees = 10, min_n = 4)
model_train(features, labels, mtry = 2, trees = 7, min_n = 4)
model_train(features, labels, mtry = 3, trees = 20, min_n = 2)
```

#### Register Model
This functionality isn't currently available in the R api. There are two ways to
approach this step:
1. Visit the tracking server and select the best performing model. Register that
model to the model registry using the tracking server UI

2. Use the `mlflow` python client to register the model. This requires knowing
the full `model_uri`, which isn't easily obtained programatically.
```{python}
import mlflow

mlflow.register_model(model_uri = 'dbfs:/mlflow/db-connect/3e00ce582037449aae17cbc441b15c72/artifacts/model', name = "iris")
```

#### Retrieve Model
The model can be retrieved from mlflow and predictions can be run against the retrieved model
```{r}
rf_model <- mlflow_load_model(model_uri = "models:/iris_r/3")
```

```{r}
mlflow_predict(rf_model, iris)
```
